### 2020-02-10 report.tex
some edits

\section{Introduction}
The vast majority of modern SAT solvers used in industrial applications are based on the conflict-driven clause learning (CDCL) algorithm. In a CDCL SAT solver, backtracking occurs after every conflict, where all literals from one or more decision levels become unassigned, before the solver resumes making decisions and performing unit propagations. It turns out that many of the literals that are unassigned during a backtrack will often be re-assigned again in roughly the same order when the solver continues. This results in many of the exact same propagation steps being repeated over and over again during the execution of the solver, and avoiding these repetitions could result in a significant speedup.

Chronological backtracking \cite{DBLP:conf/lpar/2013, DBLP:conf/sat/MohleB19, DBLP:conf/sat/NadelR18} addresses this inefficiency by heuristically backtracking fewer levels up the trail after learning a clause. However, chronological backtracking breaks some traditional invariants of the SAT solver, namely that literals are supposed to appear on the trail in monotonically increasing order by decision level, and can be complex to implement. Additionally, it does not allow the search to have as much flexibility in deciding the order of literals to assign, since leaving a large portion of the trail intact "in a sense" forces decisions originally used to construct that portion. Some empirical evidence also suggests that it is only beneficial when used sparingly (e.g., when the number of levels to be backtracked over is greater than 100).

We propose an alternative method, "trail saving", where we cache the part of the trail that is unassigned during a backtrack and then attempt to restore implied parts of it as each new decision is made. Trail saving preserves the traditional invariants of the SAT solver and its basic version is very simple to implement. It also allows the search to choose the order of decisions, but helps make propagation faster. We explore the theoretical speedup that trail saving provides, develop some enhancements to make the idea more effective, and demonstrate experimentally that it speeds up state-of-the-art SAT solvers by a non-trivial factor.

\section{Background}
We will assume the reader is familiar with SAT solving and the basics of the CDCL algorithm \cite{DBLP:series/faia/SilvaLM09}. We will use the notion of a trail to represent a sequence of literals corresponding to the current partial assignment. The literals appear on the trail in the order in which they were assigned. Associated with every literal on the trail is a decision level. Decision level 0 represents all literals which are implied by the formula itself and subsequent decisions start the beginning of a new decision level. When a literal is made true by unit propagation, the literal is given its decision level and a reason clause, where the reason clause is the clause that was made unit by propagation in order to imply this literal. Most modern solvers use the two-literal watch scheme to make unit propagation more efficient, where each clause is watched by two non-false literals. When a watcher for a clause becomes false, a replacement is searched for in the clause. If no such replacement exists

@@

	\begin{algorithm}
  	\caption{propagate()}
	\begin{algorithmic}[1]
	\WHILE{queue not empty}
	\STATE $l := $ next literal on queue
	\\BEGIN CODE TO ADD FOR TRAIL SAVING
	\IF{trail saving turned on}
		\IF{val(top of trail cache) $=$ true}
			\WHILE{trail cache not empty}
				\STATE{move top of trail cache to next literal}
				\STATE{let tc $=$ top of trail cache}
				\IF{reason(tc) is null}
					\STATE{break}
				\ELSIF{val(tc) $=$ false}
					\STATE{return conflict with reason(tc)}
				\ELSIF{val(tc) $=$ unassigned}
					\STATE{enqueue tc with reason reason(tc)}
				\ENDIF
			\ENDWHILE
		\ENDIF
	\ENDIF
	\\END CODE TO ADD FOR TRAIL SAVING
  	\FOR{each clause $c \in $ watchlist($\neg l$)}
		\FOR{each literal $x \in c$}
			\IF{$val($x$) \neq$ false and x is not already a watcher of c}
				\STATE let $x$ be a watcher of $c$ and add $c$ to watchlist($\neg x$)
				\STATE break
			\ENDIF
		\ENDFOR
		\IF{no new watcher was found}
			\IF{other watcher $x_2$ of $c = $ false}
				\STATE return c as conflict
			\ELSIF{other watcher $x_2$ of $c$ is unassigned}
				\STATE enqueue $x_2$
			\ENDIF
		\ENDIF
	\ENDFOR
	\ENDWHILE
	\end{algorithmic}
	\end{algorithm}
\clearpage

\subsection{Storing the trail cache}
The first step that is needed to use "trail saving" is to store the literals which are being unassigned during a backtrack, along with their respective reason clauses, in a "trail cache". Decisions should be marked with a null reason. The literals should appear in the same order on the trail cache as they had appeared on the trail. We will adopt the notation that the "top" of the trail cache is initially set to be the literal that appeared the earliest on the trail and will get updated as the trail cache gets used up. Note that initially the top of the trail cache is always a decison (i.e., the start of a new decision level). As the literals from the trail cache get restored (i.e., assigned), the top of the trail cache points to the next literal not yet restored.

\subsection{Restoring the trail cache}
We will base trail saving on a convenient observation. If the top of the trail cache becomes true at any point, every literal after the top of the trail cache up until the next decision literal (i.e., the next literal with a null reason) is implied by the current trail. The top of the trail cache can become true either by decision or unit propagation, and this observation still holds.\newline

\textit{Proof} Let $L_0$ be the set of literals on the trail after backtracking, $L_1$ be the set of literals on the current trail, and $L_c[i]$ denote the set of the first i literals on the trail cache. The $i$th literal on the trail cache is implied by $L_0 \cup L_c[i-1]$, as long as the $i$th literal was not a decision. $L_0 \subseteq L_1$, since $L_1$ starts out as $L_0$ and adds non-contradictory literals (by decision or unit propagation). Therefore, the $i$th literal on the trail cache is implied by $L_1 \cup L_c[i-1]$. Assume the top of the trail cache is true. Then, from above, every literal up until the next decision would be implied. $\blacksquare$\newline

Once the backtrack has been completed and the solver resumes with decisions and propagations, we will attempt to restore parts of the trail cache automatically during propagation to make it faster. When a literal is to be propagated, we first check the top of the trail cache to determine its truth value. If that literal's truth value is true, we assign all of the literals that come after it on the trail cache to be true up until (and excluding) the next the next literal with a null reason. If at any time the top of the trail cache is a literal with a non-null reason whose truth value is already false, we determine a conflict has occurred with the conflict clause being that literal's reason. At the beginning of a new backtrack, the trail cache is cleared and a new trail cache is generated.

A potentially quadratic speedup can be realized during the propagation of these literals since they are being propagated as a set, rather than one at a time. We will illustrate this with an example.\newline

With trail saving, $x2, ... x_n$ are all assigned to be true as soon as $x_1$ is made true, so immediately the values of literals $\lnot x_1, ..., \lnot x_n$ in c are false. During propagation, a new watcher is searched for to replace $\lnot x_2$, n literals are saerched over without finding any, thus c is detected as unit and $l$ becomes forced to be true. This propagation took $O(n)$ searches through clause c. $\blacksquare$\newline

Another potential speed up occurs when trail saving detects conflicts early without performing any of the pending propagations. This happens when the top of the trail cache is a literal with a non-null reason and the truth value of this literal is already false, and we can use that literal's reason clause as the conflict clause.

@@

The first enhancement to trail saving is to keep the trail cache intact instead of clearing it before each backtrack, and to add the literals that will be backtracked over to the already existing trail cache. In order to preserve soundness, we must add the literals from the current backtrack to the beginning of the trail cache, such that the trail cache has as its prefix the literals which are being backtracked over in the same order they appear on the current trail, and as its suffix the previous trail cache. In order for the suffix to be of any use, one must actually not store the last level (i.e. the conflicting level) to the trail cache at all, since the learned clause will prevent the conflicting level from being duplicated in the future. Because the trail cache can now grow indefinitely, occasionally it needs to be trimmed by removing duplicate literals. We decided to trim the trail cache whenever its size exceeded the total number of variables.

The next enhancement to trail saving is to scan the trail cache to check if there is a literal that is already assigned false within some limit. The limit we found worked best is to scan the trail until the third null reason is found (i.e. "looking ahead 2 decisions"). If a literal that is already assigned false is found, we know that we are guaranteed to have a conflict in the near future, and we force the SAT solver's subsequent decisions to come from the trail cache instead of using the regular variable selection heuristic (i.e. EVSIDS).

@@

\begin{figure}
\begin{tabular}{|l|l|l|l|l|}
\hline
                 & cadical                     & cadical\_chrono & cadical\_trail & cadical\_trail\_enhanced \\ \hline
Instances Solved & 527                         & 515             & 520            & 524                      \\ \hline
Par-2 (x $10^6$s)& 3.220                  & 3.269      & 3.248     & 3.216               \\ \hline
                 & maple                       & maple\_chrono   & maple\_trail   & maple\_trail\_enhanced          \\ \hline
Instance Solved  & 458                         & 470             & 472            & 473                      \\ \hline
Par-2 (x $10^6$s)& 3.797 & 3.690      & 3.694     & 3.667               \\ \hline
\end{tabular}
\caption{Table of results}
\end{figure}
\begin{figure}\includegraphics[scale=0.8]{cactus_cadical.pdf}\caption{\small{Comparison of run times for versions of cadical. cadical is without chronological backtracking, cadical-chrono is with chronological backtracking, cadical-trail is with the first version of trail saving, cadical-trail-2levels-ahead is the trail saving with all enhancements added.}}\end{figure}
\begin{figure}\includegraphics[scale=0.8]{cactus_maple2.pdf}\caption{\small{Comparison of run times for versions of MapleSAT. maplelcmdist is from the 2017 SAT Competition, maplelecmdist-chrono is with chronological backtracking (2018 SAT Competition), maplelcmdist-trail is with the first version of trail saving, maplelcmdist-trail-enhanced is with trail saving and all enhancements added.}}\end{figure}
\begin{figure}\includegraphics[scale=0.8]{test.pdf}\caption{}\end{figure}
\begin{figure}\includegraphics[scale=0.8]{test2.pdf}\caption{}\end{figure}
\iffalse
\begin{figure}\includegraphics[scale=0.8]{cactus_maple.pdf}\caption{\small{Comparison of run times for versions of MapleSAT. maplelcmdist is from the 2017 SAT Competition, maplelecmdist-chrono is with chronological backtracking (2018 SAT Competition), maplelcmdist-trail is with the first version of trail saving, maplelcmdist-chrono-dl is with chronological backtracking and duplicate learnts (2019 SAT Competition), maplelcmdist-trail-dl is with trail saving and duplicate learnts}}\end{figure}
\fi
\clearpage

@@

\section{Conclusion}

\bibliography{ref}{}
\bibliographystyle{splncs04}
\end{document}

>>>>

\section{Introduction}
The vast majority of modern SAT solvers that are used to solve real-world problems are based on the conflict-driven clause learning (CDCL) algorithm. In a CDCL SAT solver, backtracking occurs after every conflict, where all literals from one or more decision levels become unassigned, before the solver resumes making decisions and performing unit propagations. Traditionally, CDCL solvers would backtrack non-chronologically to the conflict level, which is the second highest decision level remaining in the conflict clause after conflict analysis has resolved away all but one literal from the current decision level \cite{DBLP:conf/dac/MoskewiczMZZM01}. Recently, however, it has been shown that backtracking chronologically (i.e. backtracking one level only after conflict analysis) \cite{DBLP:conf/lpar/2013, DBLP:conf/sat/NadelR18} can be effective on many instances and was implemented in solvers that won the last two SAT competitions. Although chronological backtracking breaks some of the conventional invariants of CDCL solvers, it has been formalized and proven correct \cite{DBLP:conf/sat/MohleB19}.

There are at least two possible side-effects of chronological backtracking which may explain its effectiveness. It has been observed that when a solver backtracks non-chronologically after a conflict, many of the literals that are unassigned during the backtrack will be re-assigned again in roughly the same order when the solver continues. By using chronological backtracking, the solver keeps a lot more of its partial assignment intact and saves from having to reconstruct a lot of the trail via propagation that it otherwise would have. Saving from having to reconstruct portions of the trail after backtracking was first introduced in the context of restarts \cite{DBLP:journals/jsat/TakRH11}. The second side-effect of chronological backtracking is that it forces the solver to stay in the same local neighbourhood during search, since it unassigns less of the partial assignment and one could argue that this might lead to finding additional conflicts more quickly or reaching a satisfying assignment (if one exists) more quickly. This second side-effect also leads to the search with chronological backtracking being less flexibile to jump out of local neighbourhoods and obey the variable selection heuristic. We argue that this second side-effect is not the primary explanation for the effectiveness of chronological backtracking, and aim to to develop a technique which will produce a benefit similar to the first side-effect while maintaining the freedom for the solver to jump out of local neighbourhoods and obey the variable selection heuristic.

Our alternative method is "trail saving", where we cache the part of the trail that is unassigned during a backtrack and then attempt to restore implied parts of it as each new decision is made. Trail saving preserves the traditional invariants of the SAT solver and its basic version is very simple to implement. It also allows the search to choose the order of decisions, but helps make propagation faster. We explore the theoretical speedup that trail saving provides, develop some enhancements to make the idea more effective, and demonstrate experimentally that it speeds up state-of-the-art SAT solvers by a non-trivial factor.

\section{Background}
We will assume the reader is familiar with SAT solving and the basics of the CDCL algorithm \cite{DBLP:series/faia/SilvaLM09}. We will use the notion of a trail to represent a sequence of literals corresponding to the current partial assignment. The literals appear on the trail in the order in which they were assigned. Associated with every literal on the trail is a decision level. Decision level 0 represents all literals which are implied by the formula itself and subsequent decisions start the beginning of a new decision level. When a literal is made true by unit propagation, the literal is given its decision level and a reason clause, where the reason clause is the clause that was made unit by propagation in order to imply this literal. Most modern solvers use the two-literal watch scheme to make unit propagation more efficient, where each clause is watched by two non-false literals. When a watcher for a clause becomes false, a replacement is searched for in the clause. If no such replacement exists and the other watcher is true or unassigned, then the clause is unit. If no such replacement exists and the other watcher is also false, then the clause is a conflict and needs to be analyzed to produce a learnt clause.

@@

\subsection{Storing the trail cache}

The first step that is needed to use "trail saving" is to store all of the literals which are being unassigned during a backtrack, along with their respective reason clauses, in a "trail cache". Decisions should be marked with a null reason. The literals should appear in the same order on the trail cache as they had appeared on the trail. We will keep a pointer, which we will refer to as the "top" of the trail cache, and initially set it to be the literal which had appeared the earliest on the trail. Note that initially the top of the trail cache always has a null reason (i.e., it was the start of a new decision level on the previous trail). As the literals from the trail cache get restored (i.e., assigned), the top of the trail cache will be updated to the next literal not yet restored.

\subsection{Restoring the trail cache}
We will base trail saving on the following critical invariant. If the top of the trail cache becomes true at any point, even after an arbitrary number of decisions and propagations have been made since the last backtrack, the set of literals on the trail cache after the top and up to but excluding the next literal with a null reason is still implied by the current trail. The top of the trail cache can become true either by decision or unit propagation, and this invariant still holds.\newline

\textit{Proof 1} We will refer to the set of literals that become unassigned during the backtrack after conflict analysis has taken place as $L_u$, and will refer to the rest of the trail that remains assigned as $L_0$. We will refer to all of the literals that get placed on the trail after $L_0$ during subsequent propagations and decisions as $L_1$. We will refer to the current trail $L_0 \cup L_1$ as $L_T$. The trail cache will contain the literals in $L_u$ ordered by their appearance on the old trail. Let $L_c[i]$ denote the set of the first i literals on the trail cache. Let $L_c[top]$ be the top of the old trail, which gets updated as parts of the trail cache get restored.

The $i$th literal on the trail cache is implied by $L_0 \cup L_c[i-1]$, as long as the $i$th literal has a non-null reason, since $L_0 \cup L_c[i-1]$ had forced $L_c[i]$ on the previous trail. $L_0 \subseteq L_T$, since $L_T = L_0 \cup L_1$. Therefore, the $i$th literal on the trail cache is also implied by $L_T \cup L_c[i-1]$.

Assume $L_c[top]$ is true, which means that $L_c[0], ..., L_C[top] \subseteq L_1 \subseteq L_T$. Then, from above, every literal on the trail cache up to but excluding the next literal with a nun-null reason would be implied. $\blacksquare$\newline

Once the backtrack has been completed and the solver resumes with decisions and propagations, we will attempt to restore parts of the trail cache automatically during propagation to make it faster by using the invariant stated above. Each time the propagation routine is about to propagate a new literal, we first check the top of the trail cache to determine its truth value. If that literal's truth value is already true, we assign all of the literals that come after it on the trail cache to be true up until (and excluding) the next literal with a null reason.

A potentially quadratic speedup can be realized during the propagation of these literals since they are being propagated as a set, rather than one at a time, as argued in \cite{DBLP:conf/sat/HickeyB19, DBLP:journals/jair/Gent13}. Restoring literals from the trail cache automatically not only leads to quicker propagation of future literals, but also means that the propagation routine has less work to do as it searches for new watchers in clauses on watchlists. All of the reasons attached to these literals that are automatically restored will now be satisfied, thus the propagation routine can skip over them rather than having to access each of them potentially multiple times to find they are unit.

If at any time the top of the trail cache is a literal $l$ with a non-null reason whose truth value is already false, we determine a conflict has occurred with the conflict clause being $l$'s reason from the trail cache. This is because we know from Proof 1 that $l$ is implied to be true at this point by the reason on the trail cache (i.e., the reason on the trail cache is unit), but $\lnot l$ was also previously assigned to be true, therefore we have a conflict. This leads to a third potential speed up, since a conflict has been detected and none of the pending propagations have to be completed. Since trail saving can sometimes save hundreds or thousands of literals at a time, the number of propagations required to detect a conflict could be very costly.

\begin{figure}\includegraphics[scale=0.8]{trail_diagram.pdf}\caption{Storing into a trail cache after backtracking.}\end{figure}

	\begin{algorithm}
  	\TitleOfAlgo{propagate()}
	\begin{algorithmic}[1]
	\WHILE{queue not empty}
	\STATE { }
	\STATE restoreTrailCache()
	\STATE { }
	\STATE $l := $ next literal on queue
  	\FOR{each clause $c \in $ watchlist($\neg l$)}
		\FOR{each literal $x \in c$}
			\IF{$val($x$) \neq$ false and x is not already a watcher of c}
				\STATE let $x$ be a watcher of $c$ and add $c$ to watchlist($\neg x$)
				\STATE break
			\ENDIF
		\ENDFOR
		\IF{no new watcher was found}
			\IF{other watcher $x_2$ of $c = $ false}
				\STATE return c as conflict
			\ELSIF{other watcher $x_2$ of $c$ is unassigned}
				\STATE enqueue $x_2$
			\ENDIF
		\ENDIF
	\ENDFOR
	\ENDWHILE
	\end{algorithmic}
	\end{algorithm}

	\begin{algorithm}
  	\TitleOfAlgo{restoreTrailCache()}
	\begin{algorithmic}[1]
	\IF{trail saving turned on}
		\STATE{let tc $=$ top of trail cache}
		\IF{val(top of trail cache) $=$ true}
			\WHILE{trail cache not empty}
				\STATE{move top of trail cache (tc) to next literal}
				\IF{reason(tc) is null}
					\STATE{break}
				\ELSIF{val(tc) $=$ false}
					\STATE{return conflict with reason(tc)}
				\ELSIF{val(tc) $=$ unassigned}
					\STATE{enqueue tc with reason reason(tc)}
				\ENDIF
			\ENDWHILE
		\ENDIF
	\ENDIF
	\end{algorithmic}
	\end{algorithm}

@@

In the standard version of trail saving outlined above, the trail cache needs to be deleted after each backtrack in order to make room for the new backtracked literals to be cached. The first enhancement to trail saving is to keep the trail cache intact instead of clearing it before each backtrack, and to add the literals that will be backtracked over to the beginning of the already existing trail cache. In order to preserve soundness, we must add the literals from the current backtrack to the beginning of the trail cache, such that the trail cache has as its prefix the literals which are being backtracked over, and as its suffix the previous trail cache. \newline

\textit{Proof 2} We will use all of the same set names as Proof 1 and also refer to the set of literals on our previous trail cache as $L_{pc}$. We concatenate the literals in $L_{pc}$ to the end of $L_c$ to get $L_{sc}$. We know that $L_{pc}$ is a valid trail cache for $L_0 \cup L_u$. If we reach the end of $L_c$ without any conflicts, we know that $L_u \subseteq L_T$. Since $L_0 \subseteq L_T$, then $L_u \cup L_0 \subseteq L_T$ and therefore $L_{pc}$ is a valid trail cache for $L_T$. $\blacksquare$\newline

In order for the previous trail cache to be of any use, one must actually never store the last level of the current trail to the trail cache at all, since a learnt clause prevents the last level from being duplicated in the future. If any part of the previous trail cache was used to construct this last level, it must also be discarded from the previous trail cache before the previous trail cache is concatenated to the end.

When concatenating trail caches together as above, it is important to note that the running trail cache can grow indefinitely. The trail cache can be trimmed periodically in the following way. Traverse down the trail cache and whenever a literal appears more than once, delete its entry in the trail cache. Whenever a literal $x$ appears whose complement $\lnot x$ has already appeared on the trail, $x$ should be deleted and everything remaining until the end of the trail cache should be deleted (in order to keep the trail cache free of conflicts).  We decided to trim the trail cache whenever its size exceeded twice the total number of variables, as that is what worked best in preliminary experiments.\newline

The next enhancement to trail saving is to scan the trail cache to check if there is a literal that is already assigned false within some limit. The limit we found worked best is to scan the trail until the third null reason is found (i.e. "looking ahead 2 decisions"). If a literal that is already assigned false is found, we know that we are guaranteed to have a conflict in the near future, and we force the SAT solver's subsequent decisions to come from the trail cache instead of using the regular variable selection heuristic.

Sometimes using trail saving to automatically restore literals with their previous reasons will lead to a literal getting a reason that is "worse" (in terms of number of literals, but one could use a different metric) than the reason it would have ended up with under normal propagation (i.e. without trail saving). The last enhancement we made to trail saving was to stop restoring literals from the old trail as soon as we hit a reason that we suspected might be "bad". For this, we kept a running mean and standard deviation of the length of all reasons used in the solver so far, and decided to stop restoring from the trail whenever we hit a reason that was greater in length than two standard deviations above the mean length. This way, when normal propagation resumes, we give it a chance to find a better reason than the one that trail saving had stored from the last descent down the trail.

Another potential enhancement one could use is to use the trail cache to try to learn a second clause automatically when the solver reaches a conflict. However, we did not find a way to implement this to make it beneficial and the number of levels we have to scan for a second conflict would have to be more than the number of levels we already look ahead on each decision (when the solver is not under conflict).

@@

\begin{figure}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
                         & Total Solved & SAT & UNSAT & Par-2 (x 10\textasciicircum{}6 s) \\ \hline
cadical                  & 527          & 315 &  212  & 3.220                             \\ \hline
cadical\_chrono          & 515          & 305 &  210  & 3.269                             \\ \hline
cadical\_trail           & 520          & 307 &  213  & 3.248                             \\ \hline
cadical\_trail\_enhanced & 524          & 315 &  209  & 3.216                             \\ \hline
\end{tabular}
\caption{Table of results for cadical or "chrono", version published in \cite{DBLP:conf/sat/MohleB19}.}
\end{figure}

\begin{figure}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
                         & Total Solved & SAT & UNSAT \\ \hline
cadical                  & 492          & 289 &  203  \\ \hline
cadical\_chrono          & 498          & 292 &  206  \\ \hline
cadical\_trail           & 493          & 290 &  203  \\ \hline
cadical\_trail\_enhanced & 496          & 292 &  204  \\ \hline
\end{tabular}
\caption{Table of results for cadical, version pulled from github as of July 30, 2020.}
\end{figure}

\begin{figure}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
                & Total Solved & SAT & UNSAT & Par-2 (x 10\textasciicircum{}6 s) \\ \hline
maple           & 458          & 259 &  199  & 3.797                             \\ \hline
maple\_chrono   & 470          & 271 &  199  & 3.690                             \\ \hline
maple\_trail    & 472          & 271 &  201  & 3.694                             \\ \hline
maple\_enhanced & 473          & 277 &  195  & 3.667                             \\ \hline
\end{tabular}
\caption{Table of results for MapleLCMDist.}
\end{figure}
\clearpage

\section{Future Work}

With more clever data structures the efficiency of restoring the trail could be improved. There are a number of parameters which can be tuned, including how often to refresh the trail, how large the cutoff point for rejecting "long" reasons should be, how long to wait to prune the trail cache if one is appending multiple trail caches together, how far to look ahead on the trail cache for conflicts (in terms of number of decisions or total number of literals). We only experimented with a few different settings of these parameters that made sense, but more careful fine-tuning of these parameters could further enhance the effectiveness of the technique. There are potentially other techniques one could use a trail cache for, including inprocessing steps that involve probing. A look down the trail cache is an incomplete but very fast way to do a probing step, as opposed to propagation. Trail saving and chronological backtracking could be combined in some way to produce even better results.
\bibliography{ref}{}
\bibliographystyle{splncs04}
\end{document}

\iffalse
\textit{Example 1} Suppose the literal $y$ can be restored from the trail cache, and its reason clause $cr$ contains the literals ${y \lnot x_1, \lnot x_2, ..., \lnot x_n}$. This means that $x1, ..., x_n$ have become true and are currently on the trail. Suppose $x_n$ is the most recent literal on the trail and  Without trail saving, the propagation routine will have to access $cr$ from the watchlist of one of the literals  $\blacksquare$\newline

With trail saving, $x2, ... x_n$ are all assigned to be true as soon as $x_1$ is made true, so immediately the values of literals $\lnot x_1, ..., \lnot x_n$ in c are false. During propagation, a new watcher is searched for to replace $\lnot x_2$, n literals are saerched over without finding any, thus c is detected as unit and $l$ becomes forced to be true. This propagation took $O(n)$ searches through clause c. $\blacksquare$\newline
\fi

\iffalse
\begin{figure}\includegraphics[scale=0.8]{cactus_cadical.pdf}\caption{\small{Comparison of run times for versions of cadical. cadical is without chronological backtracking, cadical-chrono is with chronological backtracking, cadical-trail is with the first version of trail saving, cadical-trail-2levels-ahead is the trail saving with all enhancements added.}}\end{figure}
\begin{figure}\includegraphics[scale=0.8]{cactus_maple2.pdf}\caption{\small{Comparison of run times for versions of MapleSAT. maplelcmdist is from the 2017 SAT Competition, maplelecmdist-chrono is with chronological backtracking (2018 SAT Competition), maplelcmdist-trail is with the first version of trail saving, maplelcmdist-trail-enhanced is with trail saving and all enhancements added.}}\end{figure}
\begin{figure}\includegraphics[scale=0.8]{test.pdf}\caption{}\end{figure}
\begin{figure}\includegraphics[scale=0.8]{test2.pdf}\caption{}\end{figure}
\fi
\iffalse
\begin{figure}\includegraphics[scale=0.8]{cactus_maple.pdf}\caption{\small{Comparison of run times for versions of MapleSAT. maplelcmdist is from the 2017 SAT Competition, maplelecmdist-chrono is with chronological backtracking (2018 SAT Competition), maplelcmdist-trail is with the first version of trail saving, maplelcmdist-chrono-dl is with chronological backtracking and duplicate learnts (2019 SAT Competition), maplelcmdist-trail-dl is with trail saving and duplicate learnts}}\end{figure}
\fi

@@
